\mytitle

### 2021-12-09 paper.tex
extend intro
reformulation

\newcommand{\mytitle}{An algorithm for the calculation of the real dilogarithm with instruction-level parallelism}
\newcommand{\myauthor}{Alexander Voigt}
\newcommand{\Li}{\operatorname{Li}_2}

@@

The development of modern CPUs with support for instruction\-/level
parallelism (ILP) opens the opportunity for a further reduction of the
runtime of the calculation of the dilogarithm.  To make use of ILP,
the algorithm to calculate the dilogarithm must be formulated
appropriately in order to allow for the parallel execution of multiple
floating\-/point operations.  In this paper an algorithm for the
calculation of the real dilogarithm $\Li$ is presented, which makes
use of ILP, resulting in a reduced runtime compared to the algorithms
presented so far.  A C implementation of this algorithm can be found
in the appendix and in the arXiv submission of this paper.

\section{Comparison of different implementation strategies}

As a toy example we consider the function $\ln(1+x)$, which has the
Taylor expansion
%
\begin{align}
  \ln(1+x) \approx x - \frac{x^2}{2} + \frac{x^3}{3} - \frac{x^4}{4} + \frac{x^5}{5} - \frac{x^6}{6} + \frac{x^7}{7}
  \label{eq:taylor}
\end{align}
%
around $x=0$ for $x\in(-1,1]$, where terms of $O(x^8)$ and beyond have
been omitted.  In the following we will briefly study various possible
implementations of this function and compare the performance of the
different implementations.  For brevity we will restrict ourselves to
$x\in(-1,1]$.  In C the function in Eq.~\eqref{eq:taylor} could be
implemented very naively as follows:
%
\lstinputlisting[language=C]{src/log1p_naive.c}
%
This naive implementation requires 26 floating\-/point multiplications
and 6 floating\-/point additions.\footnote{Modern compilers usually
  rewrite expensive floating\-/point divisions by a constant, such as
  $x/2$, as less expensive floating\-/point multiplication by a
  constant, for example as $0.5x$.  In the following we will therefore
  count a division of a variable by a constant as a floating\-/point
  multiplication.}

@@

\section{An ILP-aware algorithm for the real dilogarithm}

>>>>

\newcommand{\mytitle}{An algorithm for the calculation of the real dilogarithm with instruction-level parallelism}
\newcommand{\myauthor}{Alexander Voigt}
\newcommand{\Li}{\operatorname{Li}_2}
\newcommand{\secref}[1]{Section~\ref{#1}}

@@

The development of modern CPUs with support for instruction\-/level
parallelism (ILP) opens the opportunity for a further reduction of the
runtime of the calculation of the dilogarithm.  To make use of ILP,
the algorithm to calculate the dilogarithm must be formulated
appropriately in order to allow for the parallel execution of multiple
floating\-/point operations.  In this paper an algorithm for the
calculation of the real dilogarithm $\Li$ is presented, which makes
use of ILP, resulting in a reduced runtime compared to the algorithms
presented so far.

This paper is organized as follows: In \secref{sec:toymodel} different
implementation strategies are discussed using a toy example function.
In \secref{sec:algorithm} an ILP-aware algorithm for the numerical
calculation of the dilogarithm is presented.  A C implementation of
this algorithm can be found in the appendix and in the arXiv
submission of this paper.

\section{Comparison of different implementation strategies}
\label{sec:toymodel}

As a toy example we consider the function $\ln(1+x)$, which has the
Taylor expansion
%
\begin{align}
  \ln(1+x) \approx x - \frac{x^2}{2} + \frac{x^3}{3} - \frac{x^4}{4} + \frac{x^5}{5} - \frac{x^6}{6} + \frac{x^7}{7}
  \label{eq:taylor}
\end{align}
%
around $x=0$ for $x\in(-1,1]$, where terms of $O(x^8)$ and beyond have
been omitted.  In the following we will briefly show various possible
implementations of this function and discuss the performance of the
different implementations.  For brevity we will restrict ourselves to
$x\in(-1,1]$.  In C the function in Eq.~\eqref{eq:taylor} could be
implemented very naively as follows:
%
\lstinputlisting[language=C]{src/log1p_naive.c}
%
This naive implementation requires 26 floating\-/point multiplications
and 6 floating\-/point additions.\footnote{Modern compilers usually
  rewrite expensive floating\-/point divisions by a constant, such as
  $x/2$, as less expensive floating\-/point multiplication by a
  constant, for example as $0.5x$.  In the following we will therefore
  count a division of a variable by a constant as a floating\-/point
  multiplication.}

@@

\section{An ILP-aware algorithm for the real dilogarithm}
\label{sec:algorithm}
